---
title: "Home"
execute: 
  echo: false
  warnings: false
---
By using a data set from Norway on 110,204 patient admissions of people with Sepsis from 2011 to 2012, we wanted to understand the probability of survival based upon the factors of numbers of previous episodes of Sepsis and gender. The data was sourced from Kaggle on containing de-identified data on hospitalized patients. It includes the variables of sex, age, episode number and survival outcome.Though healthcare has dramatically changed over the years, we believe for these predictions to hold true. We used a logistic regression model with the variables of age (numeric variable), sex (categorical variable) and episode number (numeric) set to the outcome of the probability of survival for the i-th individual. From modeling the variables as a logistic function to the outcome, we discovered that being female can increases the odds of survival by 18%, higher age decreases the probability of survival by 4%, while higher episode numbers have slightly lower chance of survival of about 3%, though it may be as much as 5% or as low as 0%.


### Sepsis Survival Rate:
<small>(Episode number refers to the number of times a patient has had Sepsis)</small>



```{r}
#| label: Setup
#| message: false
# Load in all the libraries that may be used 
library(tidyverse)
library(gtsummary)
library(tidybayes)
library(brms)
library(broom.mixed)
# Read in the data
sepsis_data <- read.csv("data/s41598-020-73558-3_sepsis_survival_primary_cohort.csv")
```

```{r}
# Mutate and edit the data
#| label: edit data
sepsis_data <- sepsis_data |>
# Rework the sex column to make more sense
  mutate(
# Make cloumn names shorter and easier to work with
    age = age_years,
    sex = factor(sex_0male_1female, labels = c("Male", "Female")),
    outcome = hospital_outcome_1alive_0dead
  )

```


```{r}
#| label: model
#| cache: true
#| message: false
#| warnings: false
# Make the model with all covariates to outcome.
fit_sepsis <- brm(
  formula = outcome ~ age + sex + episode_number,
  family = bernoulli(link = "logit"),
  data = sepsis_data,
  silent = 2, 
  refresh = 0)

```


```{r}
#| label: make new object
# Generate new observations for prediction
newobs <- expand.grid(
  episode_number = c(1, 2, 3, 4, 5), 
  sex = c("Male", "Female"), 
  # use average age
  age = mean(sepsis_data$age)
)

```

```{r}
#| label: plot

fit_sepsis |> 
  add_epred_draws(newdata = newobs) |>
# Make the graph with x as the predicted surrvival rate, y as the episode numbers, and color as the sex
  ggplot(aes(x = .epred, y = as.factor(episode_number), fill = sex)) + 
# Chose to make stat_slab
  stat_slab(alpha = 0.5, position = "identity") + 
# Manually add colors for better customizatino
  scale_fill_manual(values = c("Male" = "cornflowerblue", "Female"  = "darkred")) + 
# Change x-axis labels to percent format
  scale_x_continuous(labels = scales::label_percent(accuracy = 1)) +
# Add labels and key
  labs(
    title = "Probability of Surviving by Episode Number and Gender",
    x = "Predicted Probability of Surviving (%)",
    y = "Episode Number",
    fill = "Sex"
  ) +
# Make a cleaner theme
  theme_minimal() +
  theme(legend.position = "bottom")

```


### Overview of Sepsis:
Sepsis is a dangerous condition that can occur to anyone. It happens when the body’s immune system responds improperly to infection and damages healthy tissue. This can lead to organ failure, widespread inflammation and possibly death. Sepsis itself is not a disease, but rather caused by other infectious diseases. 

  Important symptoms to lookout for are...
  
    -  Rashes
    
    -  Change in heart rate
    
    -  Change in body temperature
    
    -  Fatigue
    
    -  Disorientation
    
    -  Discomfort
    
    -  Clammy or sweaty skin
    
    -  Shivers or chills

  